---
title: "Final Assignment"
author: "Dulshika Dassanayake"
date: "2026-04-06"
output:
  pdf_document:
    toc: true
  md_document:
    variant: gfm
  html_document:
    df_print: paged
  word_document:
---


# 1. Notes on the recap material

## Data subsetting

### example dataset: mtcars
```{r}
mtcars
```

## without subset function

### Give the name of the data set then open square bracket[raw,]column
```{r}
mtcars [1,] # raw one only
mtcars [1,3] # raw one column 3
```

### find all rows such that vs=1 data frame $ sign vs == 1 and need to put a comma

```{r}
mtcars [mtcars$vs == 1,]
mtcars $hp[mtcars$vs == 1] # no comma allows to access all row and column
mtcars $hp[mtcars$vs == 1][1] # get the first number there
mtcars $hp[mtcars$vs == 1 & mtcars$gear == 4] # and function
max (mtcars$hp) # get the maximum value of the mtcars hp row
```

## With subset function
```{r}
subset(mtcars, vs==1)
subset(mtcars, vs==1 & gear == 4) # and
subset(mtcars, vs==1 | gear == 4) # or

```

# 2. Extra optional materials provided, the code and how it works.

### Short justification for choosing this and what want to learn

I chose the “Putting points on a map” module because spatial visualization is important for my research on 'phosphorus internal loading in lakes', where identifying geographic hotspots is essential. In this module, I explored how ggplot2 uses longitude and latitude variables to create maps by layering polygons for geographic boundaries and plotting data points using geom_point(). Learning how these functions work helps me understand how to visualize spatial environmental data clearly on maps. 


# 3. Introduction to renv. 

## Follow the coding demonstration in class. Turn in notes

### renv package

In R, packages are normally installed in a global library, meaning the same packages are shared across all projects. Because package versions may differ between computers, this can cause code to break when a project is moved to another machine.

This creates problems for reproducibility, since the same code may not produce the same results if package versions are different.

The renv package solves this problem by creating a project-specific library that records the exact versions of packages used in the project.

This ensures that the project environment can be recreated on other computers, improving reproducibility and portability of the code.

### Steps followed:

#### Create a GitHub repository
First, a GitHub repository was created to store the project files. For this demonstration, I used the already created class repository.

#### Load the necessary packages
The required packages were loaded in R to demonstrate how renv manages package versions.

```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
```

#### Initialize the renv project environment
A project-specific library and lock file were created using renv.

- Run the code in the console: renv::init()
- Record the package versions used in the project: renv::snapshot() 
- These commands create a project-specific library and a lock file '(renv.lock)'     that records the exact versions of all packages used.

#### Commit the renv files to GitHub
The renv.lock file and related project files were committed and pushed to GitHub so that the project environment can be reproduced elsewhere.

#### Recreate the environment on another computer
To run the project on a different computer:

- Download the repository from GitHub.
- Open the project in RStudio.
- Install renv if it is not already installed with: install.packages("renv")
- Restore the project environment using: renv::restore()

This installs the exact package versions listed in the renv.lock file, ensuring the project runs the same way on different machines.


                      
# GitHub link: https://github.com/Dulshikamadushani/R-Practice/tree/main/Bonus%20material

