---
title: 'Coding challenge 7: Linear model'
author: "Dulshika Dassanayake"
date: "2026-03-26"
output: 
  md_document:
    variant: gfm
  pdf_document:
    toc: true
  html_document:
    df_print: paged
  word_document:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1: Read the data
```{r}
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)

STAND <- read.csv("PlantEmergence.csv")

STAND$Treatment <- as.factor(STAND$Treatment)
STAND$DaysAfterPlanting <- as.factor(STAND$DaysAfterPlanting)
STAND$Rep <- as.factor(STAND$Rep)

str(STAND)
```

# Question 2: Fit the full linear model with interaction
```{r}
lm.full <- lm(Emergence ~ Treatment * DaysAfterPlanting, data = STAND)

summary(lm.full)
anova(lm.full)
```

## the summary of the linear model and ANOVA results:

Main results for the full model

ANOVA

Treatment: highly significant, p < 2e-16
DaysAfterPlanting: significant, p = 1.88e-05
Treatment × DaysAfterPlanting: not significant, p = 1.00

What this means

Treatment strongly affects emergence. Days after planting also affects emergence. But the interaction is not significant, so the treatment effect does not appear to change across days after planting.


# Question 3: simplified linear model without the interaction term 

Do need the interaction term?

No. Since the interaction term is not significant, should remove it and fit the simpler additive model. If the interaction is not significant, simplify the model.

```{r}
lm.simple <- lm(Emergence ~ Treatment + DaysAfterPlanting, data = STAND)

summary(lm.simple)
anova(lm.simple)
```

## Main results for the simplified model:

ANOVA

Treatment: highly significant, p < 2e-16
DaysAfterPlanting: significant, p = 1.58e-06

Important coefficients from the simplified model

Intercept = 182.1632
Treatment2 coefficient = -134.5312

Interpretation of the intercept
Because R uses the first level as the reference level, the intercept is the predicted emergence for:
Treatment 1
DaysAfterPlanting = 7

So, the expected emergence for Treatment 1 at 7 days after planting is about 182.16 plants.


Interpretation of the coefficient for Treatment 2

The coefficient for Treatment 2 is -134.53.
That means, holding DaysAfterPlanting constant, Treatment 2 has about 134.53 fewer emerged plants than Treatment 1, on average.


# Question 4: Least-square means for Treatment and Tukey grouping
```{r}
lsm.trt <- emmeans(lm.simple, ~ Treatment)
lsm.trt

cld(lsm.trt, alpha = 0.05, Letters = letters, reversed = TRUE)
```
## Interpretation of the Tukey grouping

From the treatment comparisons:

Treatment 2 is much lower than all other treatments.
Most of the other treatments are not significantly different from one another.
Treatment 5 is significantly higher than Treatment 7.
The remaining non-reference treatments mostly overlap statistically.

## Interpretation

Treatments sharing at least one letter are not significantly different.
Treatments with no letters in common are significantly different.

So:

Treatment 2 stands alone as the worst treatment.
Treatment 5 is among the highest.
Treatment 7 is lower than Treatment 5, but not clearly different from several of the other treatments.

# Question 5: Plot with letters
```{r}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor)))
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters)
  
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}
plot_cldbars_onefactor(lm.simple, "Treatment")
```

## What the letters mean

The letters show the Tukey groupings:

Bars with the same letter are not significantly different
Bars with different letters and no overlap are significantly different

# Question 6: GitHub
My GitHub link: https://github.com/Dulshikamadushani/R-Practice/tree/main/Coding%20challenge_7





















